pirouette: the error BEAST2 makes in inferring a phylogeny

### 2019-01-21 article.tex
@Giappo will love this

%Not only can phylogenies show us how species (or other
%evolutionary units) are related to each other, 
%but we can also estimate relevant parameters such as extinction and 
%speciation rates from them.
%There are many phylogenetics tools available to obtain an estimate 
%of the phylogeny of a given set of species. 
%BEAST2 (\cite{bouckaert2014beast}) is one of the most widely used ones.
%It uses a Bayesian statistical framework to estimate 
%the joint posterior distribution of phylogenies and model parameters, 
%from one or more DNA, RNA or amino acid alignments (see figure 1 
%for an overview of the workflow).

%Here, we present \verb;pirouette;:
%’BEAUti 2, BEAST2 and Tracer for R’, 
%which creates BEAST2 (v.2.4.7) configuration files,
%runs BEAST2, and analyzes its results,
%all from an R function call. This
%will save time, tedious mouse clicking and 
%reduces the chances of errors in such repetitive actions.
%The interface of \verb;pirouette; mimics the tools it
%is based on. This
%familiarity helps both beginner and experienced BEAST2 users 
%to make the step from those tools to \verb;pirouette;.
%\verb;pirouette; enables the creation of a single-script 
%pipeline from sequence alignments to posterior analysis in R.

@@

The fourth step measures the inference error, using the phylogenies in the
Bayesian posterior.

From the
DNA alignment a Bayesian posterior is generated, using a user-specified
inference setup. The given tree is compared with all posterior trees, using
a user-specified error measurement setup.

A first and optional step is to generate a 'twin tree', that will be
analyzed in the same way as the true tree.

and enables the full BEAST2 workflow from a single R function call,
%in a similar way to what subsequent usage of BEAUti, DensiTree and Tracer would produce.
%\verb;pirouette;'s main function is \verb;bbt_run;, which
%configures BEAST2, runs it and parses its output. 
%\verb;bbt_run; needs at least the name of a 
%FASTA file containing a DNA alignment. 
%The default settings for the other arguments of \verb;bbt_run; 
%are identical to BEAUti's and BEAST2's default settings.
%Per alignment, a site model, clock model and tree prior can be chosen.
%Multiple alignments can be used, each with its own (unlinked) site model, 
%clock model and tree prior.

%\verb;pirouette; currently has 108 exported functions to set up  
%a BEAST2 configuration file. 
%\verb;pirouette; can currently handle the majority of BEAUti use cases.
%Because of BEAUti's high number of plugins, 
%\verb;pirouette; uses a software architecture that is designed to be extended.
%Furthermore, \verb;pirouette; has 13 exported functions to run and help run BEAST2.
%One function is used to run BEAST2, another
%one installs BEAST2 to a default location.
%Finally, \verb;pirouette; has 21 exported function to parse the BEAST2 output
%files and analyze the created posterior. \verb;pirouette; gives the
%same ESSes and summary statistics as Tracer. The data is formatted
%such that it can easily be visualized using \verb;ggplot2; (for a trace,
%similar to Tracer) or \verb;phangorn; (\cite{phangorn}) (for 
%the phylogenies in a posterior, similar to DensiTree).

%Currently, \verb;pirouette; does not contain all functionality 
%in BEAUti, BEAST2 and their many plug-ins, 
%because these tools themselves also change in time.
%\verb;pirouette; currently works only on DNA data,
%because this is the most common use case. 
%Nevertheless, \verb;pirouette; provides the majority of default tree priors and supports the most important command-line arguments of BEAST2, provides
%the core Tracer analysis options, and has the most basic subset of 
%plotting options of DensiTree. Up till now, the \verb;pirouette; features 
%implemented are those requested by users. Further extension of \verb;pirouette; 
%will be based on future user requests.

>>>>

@@

The fourth step measures the inference error, using the phylogenies in the
Bayesian posterior. These phylogenies are compared to the given
phylogeny using an error statistic, which is the nLTT 
statistic [Janzen et al., ?2016]. Additionally, the user can specify the
propertion of posterior phylogenies to discard (also called: the burn-in).

An optional step is to generate a 'twin tree', that will be
analyzed in the same way as the true tree.
